Staged Abstract Interpreters

### 2018-10-29 paper/20_bg.tex
continue working on outline
revise on intro
worked on paper
misc

\subsection{The Language}

@@

>>>>

\subsection{The Language} \label{bg_lang}

syntax, higher-order functional, variables, lambda, app, fix, let, arith, if0

concrete interpreter.

@@

big-step abstract interpreter using delimited control operators \cite{Wei:2018:RAA:3243631.3236800}.

monadic big-step abstract interpreter \cite{DBLP:journals/pacmpl/DaraisLNH17}.

\subsection{Multi-Stage Programming}

general idea; LMS framework in scala. MetaOCaml?

Other notable implementations of MSP exist in ML family, for examples, MetaML \cite{DBLP:conf/pepm/TahaS97} 
and MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14}.
Compared with the LMS approach in Scala, MetaML/MetaOCaml use term-level annotations such as brackets, 
escape, and run. Notwithstanding, we use LMS in this paper, the idea of staging an abstract interpreter
still applies with other MSP implementations.

### 2018-10-29 paper/50_cases.tex
misc

\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation as a new
implementation technique for abstract interpretation based static analysis \cite{Boucher:1996:ACN:647473.727587}.
The idea is inspired by partial evaluation and similar with the presented paper -- the program can be known 
at compile-time therefore the overhead if interpretation can be eliminated. The compiled analysis
can be represented by text or closures (higher-order functions).
Specifically, \citeauthor{Boucher:1996:ACN:647473.727587} show how to compile a monovariant control-flow 
analysis \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007} 
for continuation-passing style (CPS) programs. Since the analyzed program is in CPS, the analyzer 
is essentially a big-step CE (\textit{Control-Environment}) abstract interpreter.
After specializing the analyzer with respect to the input program, the \textit{control} part of 
abstract interpreter is eliminated, instead, the compiled analysis is a closure which takes an 
environment as argument. Therefore the overhead of traversing the abstract syntax tree of input 
program also has been eliminated.

>>>>

\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation as a new
implementation technique for abstract interpretation based static analysis \cite{Boucher:1996:ACN:647473.727587}.
The idea is inspired by partial evaluation and similar with the presented paper -- the program can be known 
at compile-time therefore the overhead if interpretation can be eliminated. The compiled analysis
can be represented by text or closures (higher-order functions).
Specifically, \citeauthor{Boucher:1996:ACN:647473.727587} show how to compile a monovariant control-flow 
analysis \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007} 
for continuation-passing style (CPS) programs. Since the analyzed program is in CPS, the analyzer 
is essentially a big-step CE (\textit{Control-Environment}) abstract interpreter.
After specializing the analyzer with respect to the input program, the \textit{control} component of 
abstract interpreter is eliminated, instead, the compiled analysis is a closure which takes an 
environment as argument. Therefore the overhead of traversing the abstract syntax tree of input 
program also has been eliminated.

### 2018-10-29 paper/70_related.tex
misc

\textbf{Abstract Compilation}. The idea in this paper is closely inspired by 
abstract compilation \cite{Boucher:1996:ACN:647473.727587}.
\citeauthor{Boucher:1996:ACN:647473.727587} presented abstract compilation
techniques as an efficient implementation of the monovariant flow analysis 
\textit{0}-CFA for programs in continuation-passing style. The key idea is to 
reduce the interpretation overhead on traversing the syntax tree by partial evaluation.
Specifically, they proposed two kinds of abstract compilers to achieve this. 
The first one is to generate specialized analysis as a textual program, then 
can be loaded and executed by \texttt{eval} or similar mechanisms. The second 
one is using closure as a representation of specialized analysis. 
A closure is a function that remembers the environment where it defined. So the 
abstract compiler generates higer-order functions on-the-fly rather than textual 
program, then the functions can be applied immediately in a higher-order host language.

\citeauthor{Johnson:2013:OAA:2500365.2500604} in their work also adopt the 
idea closure generation for optimizing small-step abstracting abstract machine
\cite{Johnson:2013:OAA:2500365.2500604}. The source program is firstly compiled
to intermediate form called "abstract bytecode", which are actually higher-order functions, 
and then executed on a abstract abstract machine.

Abstract compilation for analyzing logic programs\cite{DBLP:conf/flops/AmatoS01}

@@

\citeauthor{hardekopf2011flow} apply staging to flow-sensitive pointer analysis \cite{hardekopf2011flow}. 
They firstly stage the program code to obtain a sparse representation, then conduct
the flow-sensitive analysis.

@@

\textbf{Multi-Stage Programming}.

Collapsing Towers of Interpreters\cite{Amin:2017:CTI:3177123.3158140}

TODO: compare to other staging approach (e.g., MetaOCaml)

>>>>

\textbf{Abstract Compilation}. The idea in this paper is closely inspired by 
abstract compilation \cite{Boucher:1996:ACN:647473.727587}.
\citeauthor{Boucher:1996:ACN:647473.727587} presented abstract compilation
techniques as an efficient implementation of the monovariant flow analysis 
(\textit{0}-CFA) for programs written in continuation-passing style. 
The key idea is to remove the interpretation overhead on traversing the syntax 
tree by partial evaluation. Specifically, they proposed two similar kinds of abstract compilation techniques.
The first one is to generate specialized analysis as a textual program, then 
can be loaded and executed by \texttt{eval} or other similar mechanisms. 
The second one is using closure, i.e., a function that remembers its environment, 
as a representation of specialized analysis. 
So compiling the analysis generates higer-order functions on-the-fly with respect to the analyzed program,
then the generated closure can be applied immediately in the higher-order host language.

\citeauthor{Johnson:2013:OAA:2500365.2500604} adapt the idea closure generation 
for optimizing small-step abstract interpreter in state-transition style \cite{Johnson:2013:OAA:2500365.2500604}.
The analyzed program is firstly compiled to a intermediate representation called 
"abstract bytecode", which are actually higher-order functions, and then be executed 
on a abstract abstract machine for that IR.

%\citeauthor{DBLP:conf/flops/AmatoS01} apply abstract compilation for non par-sharing and freeness analysis in logic programs \cite{DBLP:conf/flops/AmatoS01}.

@@

\citeauthor{DBLP:conf/cgo/HardekopfL11} apply staging to flow-sensitive pointer analysis \cite{DBLP:conf/cgo/HardekopfL11}. 
They firstly stage the program code to obtain a sparse representation, then conduct
the flow-sensitive analysis.

@@

\textbf{Meta-Programming}.

Partial evaluation. 
Reflexive language \cite{DBLP:conf/gpce/Asai14}
Collapsing Towers of Interpreters \cite{Amin:2017:CTI:3177123.3158140}

Multi-stage programming. MetaML, MetaOCaml, Mint, LMS
